\LARGE \bf Artificial Cognition for Social Human-Robot Interaction:\\ An Implementation

### 2016-06-20 rebuttal2.tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editors,

\vspace{2em}

Thank you for coming back to us, we are glad that the special issue is still
going on.

We have hopefully addressed the remaining concerns raised during the second
round of reviews.

\subsection*{Language}

Your first and most stringent comment relates to the language. We have carefully
proof-read the manuscript with the help of a native English speaker.
Specifically:

\begin{itemize}
    \item a special care has been given to rephrase colloquial expressions and
        non-academic phrasing,
    \item the abstract has been entierly rewritten,
    \item all the specific language changes pointed out by the editors and
        reviewers have been integrated.
\end{itemize}

To facilitate the review process, we attach an annotated version of the
manuscript where all these changes have been highlighted.

\subsection*{Relation to the state of the art}

We initially made the choice to discuss the existing litterature in each of the
``cognitive skills'' sections, for readability reasons. We however acknowledge
that a higher-level discussion of our work with respect to the state of the art
of human-robot interaction systems would be useful.

To this end, we have split Section 5 into a discussion (Section 5) and a
conclusion (Section 6) that states clearly our contribution (including
references to all our previous publications -- Section 6.1), followed by a
comparative discussion with the related approaches (Section 6.2).

\subsection*{Reviewer Comments}

Reviewer 2's comments have all been addressed in the manuscript.
We highlight hereafter the main changes.

\begin{itemize}
    \item examples have been added to Section 2.2 to make it easier to follow,
    \item we have clarified how the system manages discrepancies between OpenCyc
        terminology and human natural language by explicit manual labelling
        (Section 2.2.2)
    \item We have precised that our approach of transient geometric situation assessment
        prevents reasoning on the situation history (Section 3.2)
    \item We have clarified how the {\tt nextTo} spatial relation is computed.
    \item We have clarified how primitive action recognition is used in
        combination with situation assessment and task planning to monitor the
        human activities (Section 3.2.3)
    \item The implicit aspect of multi-modality during dialogue processing has
        been made clearer (Section 3.3.2)
\end{itemize}

Finally, as suggested by Reviewer 3, we have changed the title of the article
from ``Human-Robot Interaction: Tackling the AI Challenges'' to the more
specific ``Artificial Cognition for Social Human-Robot Interaction: An
Implementation''.

\bibliographystyle{abbrv}
\bibliography{laas_hri}

\end{document}

### 2016-06-29 paper.tex
Update on Overleaf.

We identify the needed individual and collaborative cognitive
skills: geometric reasoning and situation assessment based on perspective-taking
and affordance analysis; acquisition and representation of knowledge models for
multiple interactants (humans and robots); natural multi-modal dialogue;
human-aware task planning; human-robot joint task achievement.  The article
discusses each of these abilities, presents a working implementation, and shows
how they combine in a coherent and original deliberative architecture for
human-robot interaction. Supported by experimental results, we eventually show
how explicit knowledge management, both symbolic and geometric, proves to be
instrumental to richer and more natural human-robot interactions by pushing for
pervasive, human-level semantics within the robot's deliberative system.

@@

We focus on a specific class of interactions: collaborative task achievement
supported by multi-modal and situated communication. Figure~\ref{fig:hri-dec}
illustrates this context: the two agents share a common space and exchange
information through multiple modalities (we specifically consider verbal
communication, deictic gestures and social gaze), and the robot is expected to achieve
interactive object manipulation, fetch and carry tasks and other similar
chores by taking into account, at every stage, the intentions, beliefs,
perspectives, skills of the human partner.  Namely, the robot must be able to
recognise, understand and participate in communication situations, both explicit
(\eg the human addresses verbally the robot) and implicit (\eg the human points
to an object); the robot must be able to take part in joint actions, both
pro-actively (by planning and proposing resulting plans to the human) and
reactively; the robot must be able to move and act in a safe, efficient and
legible way, taking into account social rules like proxemics.

@@

These three challenges, \emph{communication}, \emph{joint actions},
\emph{human-aware execution}, structure the research in human-robot
interaction. They can be understood in terms of cognitive skills that they
mandate. \emph{Joint action}, for instance, builds from:

@@

    \item a belief state that includes {\it a priori} common-sense knowledge and
        mental models of each of the interactors.

@@

\subsection{Article Overview}

The remaining of the article discusses the robotic architecture that we have
built to approach social and autonomous human-robot interaction, and how it
relates to other approaches. We organise this discussion in four sections.

The next section introduces the architecture as a whole, as well as the
knowledge model that we have developed for our robots. Then,
Section~\ref{sec:impl} details each of the main components of the architecture,
and highlights their significance for Artificial Intelligence.
Section~\ref{sec:expe} presents two studies that illustrate in a practical way
what can be currently achieved with our robots.  The
Sections~\ref{sect|discussion} and~\ref{sect|conclusion} finally restate our
main contributions and discuss the key challenges that human-robot interaction
brings to Artificial Intelligence.

@@

Bridging multiple independent software modules in one coherent robotic
architecture is not only a technical challenge, but also a design and
architectural challenge. In particular, properly managing the rich semantics of
human-level natural interaction raises a range of issues. Our basic assumption
and guiding principle is that human-level interaction is easier to achieve if
the robot itself relies internally on human-level semantics.  We implement this
principle by extensively relying on explicit knowledge representation and
manipulation: software components communicate with each other using first-order
logic statements organized into ontologies and whose semantics are close to the
ones manipulated by humans.

Figure~\ref{fig|archi} gives an overview of our architecture. An active
knowledge base ({\sc Oro}), conveniently thought as a semantic
blackboard, connects most of the modules together: the geometric
reasoning module ({\sc Spark}) produces symbolic assertions (like
\stmt{BOOK1 isOn TABLE}) describing the state and dynamics of the robot's
environment. These logical statements are stored in the knowledge base, and
queried back by the language processing module ({\sc Dialogs}), the symbolic task
planner (HATP) and the execution controller. The output of the language
processing module and the activities started by the robot controller are
likewise stored as symbolic statements.

@@

    \node[fill=Purple!50] (thing) {\textbf{thing}};
    \node [below=1.5 of thing] (ptt) {partially tangible thing} edge[dashed] (thing);
        \node [left=of ptt] {physical support} edge (ptt);
        \node [below left=of ptt] {visual mark} edge (ptt);
        \node [below right=of ptt] {container} edge (ptt);
        \node [right=of ptt] {fluid} edge (ptt);
        \node [below=1.5 of ptt] (stt) {solid tangible thing} edge (ptt);
            \node [below=1 of stt] {graspable object} edge (stt);
            \node [below left=of stt] {body part} edge (stt);
            \node [left=of stt] {furniture} edge (stt);
            \node [below right=of stt] (ea) {embodied agent} edge (stt);
                \node [below=of ea] {human} edge (ea);
                \node [right=of ea] {robot} edge (ea);

@@

\bibliographystyle{elsarticle-num}
%\bibliographystyle{abbrv}
\bibliography{laas_hri}

>>>>

We identify the needed individual and collaborative cognitive
skills: geometric reasoning and situation assessment based on perspective-taking
and affordance analysis; acquisition and representation of knowledge models for
multiple agents (humans and robots, with their specificities) natural multi-modal dialogue;
human-aware task planning; human-robot joint task achievement.  The article
discusses each of these abilities, presents a working implementation, and shows
how they combine in a coherent and original deliberative architecture for
human-robot interaction. Supported by experimental results, we eventually show
how explicit knowledge management, both symbolic and geometric, proves to be
instrumental to richer and more natural human-robot interactions by pushing for
pervasive, human-level semantics within the robot's deliberative system.

@@

The proposed scheme is specially designed for  interaction, and more particularly cooperative task achievement by one robot and one or several humans - called here the robot human partners. It is applicable to domains such as assistant or service robots in domestic and public places as well as to teammate robots in the factory of the future.

We focus on a specific class of interactions: human-robot collaborative task achievement
supported by multi-modal and situated communication. Figure~\ref{fig:hri-dec}
illustrates this context: the human and the robot share a common space and exchange
information through multiple modalities (we specifically consider verbal
communication, deictic gestures and social gaze), and the robot is expected to achieve
interactive object manipulation, fetch and carry tasks and other similar
chores by taking into account, at every stage, the intentions, beliefs,
perspectives, skills of its human partner.  Namely, the robot must be able to
recognise, understand and participate in communication situations, both explicit
(\eg the human addresses verbally the robot) and implicit (\eg the human points
to an object); the robot must be able to take part in joint actions, both
pro-actively (by planning and proposing resulting plans to the human) and
reactively; the robot must be able to move and act in a safe, efficient and
legible way, taking into account social rules like proxemics.

@@

These three challenges, \emph{communication}, \emph{joint action},
\emph{human-aware execution}, structure the research in human-robot
interaction. They can be understood in terms of cognitive skills that they
mandate. \emph{Joint action}, for instance, builds from:

@@

    \item a belief state that includes {\it a priori} common-sense knowledge and
        mental models of each of the agents involved (the robot and its human partners).

@@

\subsection{Contribution and Article Overview}

The main contributions concern the architecture of the decisional layer of a robot designed to share 
the space and the task with humans and to act and interact in a way that facilitates human action and decision.

We will show in the sequel that:
\begin{itemize}

\item such architecture allows to achieve advanced grounding in a relatively complex environment 
involving one or several humans and a robot

\item  it is able, thanks to the combination of perspective taking and affordances computation 
and with the help of an inference engine, to provide pertinent knowledge for situated dialog

\item  this architecture provides a framework and and a set of mechanisms that allow 
the robot to reason about the mental state of its human partners

\item  the same set of affordances and induced facts together with an explicit reasoning
on humans and robot abilities in a given context,  allow to synthesize relevant human-robot shared plans

\item  it opens the possibility and provides a propitious context for human-aware planning 
at symbolic as well as at geometric level

Another important aspect of such an architecture is that it has been fully implemented and 
demonstrated on several use cases. It is also noteworthy that it proved to be a pertinent 
framework  for new contributions about human-robot joint  action as well as for multi-disciplinary studies.

While several contributions in the literature provide very interesting insights and contributions 
on one aspect or another, we are not aware of a fully implemented architecture that effectively
combines in a coherent way all the issues mentioned above. Besides, this has been done with an effort
to make explicit all the knowledge involved.

We do not claim that this is the sole architecture that can be designed for such problems nor that it
covers all the issues. Our point here is that it is a fully implemented architecture that allowed us to
identify, refine and combine a number of key cognitive abilities that we believe are necessary. Indeed, 
we show here below that it provided us a mean to develop and explore a number of key features for HRI 
in terms of models and algorithms and their articulation in an integrated decisional system for a 
cognitive and interactive robot.

\vspace{0.5cm}
The remaining of the article discusses the robotic architecture that we have
built to address social and autonomous human-robot interaction, and how it
relates to other approaches. We organise this discussion in four sections.

The next section introduces the architecture as a whole, as well as the
knowledge model that we have developed for our robots. Then,
Section~\ref{sec:impl} provides further details on each of the main components of the architecture.
Section~\ref{sec:expe} presents two studies that illustrate in a practical way
what can be currently achieved with our robots.  The
Sections~\ref{sect|discussion} and~\ref{sect|conclusion} finally summarize our
main contributions and discuss the key challenges that human-robot interaction
brings to Artificial Intelligence.

@@

Articulating multiple independent software modules in one coherent robotic
architecture is not only a technical challenge, but also a design and
architectural challenge. In particular, properly managing the rich semantics of
human-level natural interaction raises a range of issues. Our basic assumption
and guiding principle is that human-level interaction is easier to achieve if
the robot itself relies internally on human-level semantics.  We implement this
principle by extensively relying on explicit knowledge representation and
manipulation: software components communicate with each other using first-order
logic statements organized into ontologies and whose semantics are close to the
ones manipulated by humans.

Figure~\ref{fig|archi} gives an overview of our architecture. An active
knowledge base ({\sc Oro}), conveniently thought as a semantic
blackboard, connects most of the modules: the geometric
reasoning module ({\sc Spark}) produces ar relatively high frequency symbolic assertions (like
\stmt{BOOK1 isOn TABLE}) describing the state of the robot environment and its evolution over time. 
These logical statements are stored in the knowledge base, and
queried back, when necessary, by the language processing module ({\sc Dialogs}), the symbolic task
planner (HATP) and the execution controller. The output of the language
processing module and the activities managed by the robot controller are
likewise stored as symbolic statements.

@@

    \node[fill=Purple!50] (thing) {\textbf{thing}};
    \node [below=1.5 of thing] (ptt) {partially tangible thing} edge[dashed] (thing);
        \node [left=of ptt] {physical support} edge (ptt);
        \node [below left=of ptt] {visual mark} edge (ptt);
        \node [below right=of ptt] {container} edge (ptt);
        \node [right=of ptt] {fluid} edge (ptt);
        \node [below=1.5 of ptt] (stt) {solid tangible thing} edge (ptt);
            \node [below=1 of stt] {graspable object} edge (stt);
            \node [below left=of stt] {body part} edge (stt);
            \node [left=of stt] {furniture} edge (stt);
            \node [below right=of stt] (ea) {cited agent} edge (stt);
                \node [below=of ea] {human} edge (ea);
                \node [right=of ea] {robot} edge (ea);

@@

\bibliographystyle{elsarticle-num}
%\bibliographystyle{abbrv}
\bibliography{laas_hri,other_citations}
